---
title: "Lower Basin Projected Water Use: 2025 - 2027"
subtitle: "`r ifelse(params$scenario == 'Most', paste0(params$mon_year, ': ', params$scenario, ' Probable 24-Month Study'), paste0(params$mon_year, ': Probable ', params$scenario, ' 24-Month Study'))`"
format:
  docx:
    mainfont: Arial # Sets the main font to Arial for Word output
    reference-doc: custom-reference-doc.docx # Use template
filters:
  - pandoc-word-styles.lua
params:
  scenario: 
    label: "Select Scenario"
    value: "Most"  # Change this to the required scenario
  mon_year:
    label: "24MS Month and Year"
    value: "April 2025" # Change this to the required month and Year
  proj_state_use:
    label: "Projected State Use file path (relative or absolute)"
    value: "data/Projected State Use -APR25.xlsx"
  most_rng:
    label: "Range of the MOST data in Proj. State Use"
    value: "A1:D58"
  min_rng:
    label: "Range of the MIN data in Proj. State Use"
    value: "F1:I58"
  max_rng:
    label: "Range of the MAX data in Proj. State Use"
    value: "K1:N58"
  sys_cons_sht:
    label: "Sheet name of System Cons data in Proj. State Use file"
    value: "System Conservation"
  cons_CA_rng:
    label: "CA Conservation data range"
    value: "AH4:AM20"
  cons_nonCAWCD_rng:
    label: "non-CAWCD Conservation data range"
    value: "Z4:AE20"
  cons_CAWCD_rng:
    label: "CAWCD Conservation data range"
    value: "R4:W20"
  cons_summary_sht:
    label: "Conservation Summary Table sheet name"
    value: "SysCon SummaryTable"
  cons_summary_rng:
    label: "Range of the Conservation Summary Table"
    value: "C79:I103"
  sct_data:
    label: "SCT data output file path (relative or absolute)"
    value: "data/APR25-Most.xlsx"
  sct_data_sht:
    label: "Sheetname of the SCT data"
    value: "Sheet1"
  sct_data_rng:
    label: "Range of the SCT data"
    value: "A1:E55"

    
---

```{r setup, include=FALSE}
# This chunk is for global R options and loading packages
# Set echo=FALSE to hide code, warning=FALSE to hide warnings, message=FALSE to hide messages
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# packages_needed <- c("tidyverse", "knitr", "readxl", "janitor", "stringr",
#                      "flextable", "officer", "zoo")

# # Loop through the package list and install if not already installed
# for (pkg in packages_needed) {
#   if (!requireNamespace(pkg, quietly = TRUE)) {
#     install.packages(pkg)
#   }
# }

library(tidyverse) # To use tidyverse functions
library(knitr)     # For kable tables
library(readxl)
library(janitor)
library(stringr)
library(flextable)
library(officer)
library(zoo)

##### Update file paths ######################
## File path for Projected State Use file
file_path <- params$proj_state_use

# file path for the Water Use Assumption SCT export excel file
# Update sheet name and data range for the Water Use Assumption SCT export
file_path2 <- params$sct_data

sheet_name4 <- params$sct_data_sht
data_range4 <- params$sct_data_rng

##############################################


## These sheet names and data ranges (for Projected State Use) should not be changed.
sheet_name1 <- as.character(year(Sys.Date()))
sheet_name2 <- as.character(year(Sys.Date())+1)
sheet_name3 <- as.character(year(Sys.Date())+2)
data_range_most <- params$most_rng # MOST Data range
data_range_min <- params$min_rng # MIN Data range
data_range_max <- params$max_rng # MAX Data range
# Conservation data ranges
rangeCA = params$cons_CA_rng
rangeNonCAWCD = params$cons_nonCAWCD_rng
rangeCAWCD = params$cons_CAWCD_rng

## System Conservation sheet name ####
sheet_cons = params$sys_cons_sht

### update conservation table sheet name and table range ######
sheet_cons_table = params$cons_summary_sht
cons_table_range = params$cons_summary_rng

#####################################################################
# Access the parameter value
selected_scenario <- params$scenario
source("helper-functions_v1.R")

# conditional data loading 
if (selected_scenario == "Most") {
  Most1 <- process_excel_data(file_path, sheet_name1, data_range_most)
  Most2 <- process_excel_data(file_path, sheet_name2, data_range_most)
  Most3 <- process_excel_data(file_path, sheet_name3, data_range_most)
  print("Processed 'Most' scenario data.")
} else if (selected_scenario == "Min") {
  Min1 <- process_excel_data(file_path, sheet_name1, data_range_min)
  Min2 <- process_excel_data(file_path, sheet_name2, data_range_min)
  Min3 <- process_excel_data(file_path, sheet_name3, data_range_min)
  print("Processed 'Min' scenario data.")
} else if (selected_scenario == "Max") {
  Max1 <- process_excel_data(file_path, sheet_name1, data_range_max)
  Max2 <- process_excel_data(file_path, sheet_name2, data_range_max)
  Max3 <- process_excel_data(file_path, sheet_name3, data_range_max)
  print("Processed 'Max' scenario data.")
} else {
  stop("Invalid scenario selected. Please choose 'Most', 'Min', or 'Max'.")
}

# create generic names for the dataframes based on the selected scenario.
if (selected_scenario == "Most") {
  DataY1  <- Most1
  DataY2 <- Most2
  DataY3<- Most3
} else if (selected_scenario == "Min") {
  DataY1  <- Min1
  DataY2 <- Min2
  DataY3<- Min3
} else if (selected_scenario == "Max") {
  DataY1  <- Max1
  DataY2 <- Max2
  DataY3<- Max3
}

```

### `r current_year`

Total projected water use **(`r sprintf("%.3f",round(as.numeric(find_rows_containing_string(DataY1,"Total Basin Use")$value)/1000000,3))` maf)** – `<w:r><w:rPr><w:highlight w:val="yellow"/><w:i/></w:rPr><w:t>Based on Lake Mead Operating Condition of Level 1 Shortage and water savings contributions under the LB DCP Agreement and IBWC Minute 323.</w:t></w:r>`{=openxml}

#### U.S. Contractors: `r sprintf("%.3f",round(sum(apportionment[,2])/1000000,3))` maf

##### California: [`r sprintf("%.3f",round(sum(apportionment[1,2])/1000000,3))` maf]{style="color: #0070C0;"}
* MWD annual diversion of `r prettyNum(round(as.numeric(find_rows_containing_string(sct_data,"MWDDiversionAnnualFC")[2])/1000,0),  big.mark = ",", scientific = FALSE)` kaf
    
    
```{r nwd-diversion-y1, results='asis'}
cat("  * Projected diversion includes the ")
cat(mwdICSY1)
cat(" kaf of ICS")

if(find_rows_containing_string(sct_data, "AnnualDCPContribution_Coachella")[1,2]>0){
  addCADCP = TRUE
  coachDCP = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDCPContribution_Coachella")[1,2]/1000,1)))
  mwdDCP = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDCPContribution_MWD")[1,2]/1000,1)))
} else{
  addCADCP = FALSE
  coachDCP=""
  mwdDCP=""
}

if (exists("addCADCP") && addCADCP) {
  
  cat(" and ")
  cat(coachDCP)
  cat(" kaf of water stored for CVWD's share of California's DCP contribution\n")
  cat("* DCP contribution of ")
  cat(mwdDCP + coachDCP)
  cat(" kaf through EC ICS conversion")
}

cat("\n")

```

```{r generate-bullet-list-apply, results='asis'}
# Check if cons_CA_CY has rows and at least two columns before proceeding
if (nrow(cons_CA_CY) > 0 && ncol(cons_CA_CY) >= 2) {
  cat("* Total California System Conservation of ")
  cat(sprintf("%.1f",round(sum(cons_CA[,column_number])/1000,1)))
  cat(" kaf\n")
  # Use apply to iterate over rows. For each row:
  # row[1] will be the value from the first column of cons_CA_CY
  # row[2] will be the value from the second column of cons_CA_CY
  bullet_points <- apply(cons_CA_CY, 1, function(row) {
    paste0("  * ", row[1], " system conservation of ", round(as.numeric(row[2])/1000,1) , " kaf\n")
  })
  
  # Print each bullet point, separated by a newline
  cat(paste0(bullet_points, collapse = ""))
  
}

```

* Needles PSCP volume of `r as.numeric(find_rows_containing_string(DataY1, "Needles")$value)*-1` af
```{r conditional-CA-BICS-Y1, results='asis', echo=FALSE}

if(find_rows_containing_string(sct_data, "AnnualCreationBiNat_MWD")[1,2]>0){
  addBICS = TRUE
  bics = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationBiNat_MWD")[1,2]/1000,1)))
} else{
  addBICS = FALSE
  bics=""
}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by MWD and IID\n")
}


```

##### Arizona: [`r sprintf("%.3f",round(sum(apportionment[2,2])/1000000,3))` maf]{style="color: #0070C0;"}
* CAP annual diversion of `r prettyNum(round(find_rows_containing_string(sct_data, "CAPAnnualFC")[1,2]/1000,0),  big.mark = ",", scientific = FALSE)` kaf

```{r conditional-AZ-reductions-Y1, results='asis', echo=FALSE}

addreductions = calculate_AZ_reductions(year=1, sct_data = sct_data)$addreductions
reductions_scen_all = calculate_AZ_reductions(year=1, sct_data = sct_data)$reductions_scen_all

if (exists("addreductions") && addreductions) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("    * Projected diversion includes a ")
  cat(reductions_scen_all)
  cat("\n")
}

az_dcpICS <- find_rows_containing_string(sct_data, "AnnualCreationDCP_CAWCD")[2]
az_dcpICS <- ifelse((az_dcpICS/1000) %% 1 != 0, round(az_dcpICS/1000, 1), round(az_dcpICS/1000, 0))
az_dcpSys <- find_rows_containing_string(sct_data, "AnnualSysWaterforDCP_CAWCD")[2]
az_dcpSys <- ifelse((az_dcpSys/1000) %% 1 != 0, round(az_dcpSys/1000, 1), round(az_dcpSys/1000, 0))

```

* DCP contribution will be made by creating `r ifelse(az_dcpICS > 0, paste0(az_dcpICS, " kaf of ICS and ", az_dcpSys, " kaf of non-ICS water"), paste0(az_dcpSys, " kaf of non-ICS water"))` 
* Total non-CAWCD System Conservation of `r sprintf("%.1f",round(sum(cons_NonCAWCD_filt[,column_number])/1000,1))` kaf
```{r generate-bullet-list-nonCAWCD-y1, results='asis'}
# Check if cons_CA_CY has rows and at least two columns before proceeding
if (nrow(cons_NonCAWCD_CY) > 0 && ncol(cons_NonCAWCD_CY) >= 2) {

  # Use apply to iterate over rows. For each row:
  # row[1] will be the value from the first column of cons_CA_CY
  # row[2] will be the value from the second column of cons_CA_CY
  bullet_points <- apply(cons_NonCAWCD_CY, 1, function(row) {
    paste0("  * ", row[1], " : ", ifelse(as.numeric(row[2])>100, paste0(round(((as.numeric(row[2])/1000)*100+.1),-1)/100 , " kaf\n"), paste0(round((as.numeric(row[2])*10+.1),-1)/10, " af\n") ))
  })
  
  # Print each bullet point, separated by a newline
  cat(paste0(bullet_points, collapse = ""))
  
} else if (nrow(cons_NonCAWCD_CY) == 0) {
  cat("No data available in `cons_NonCAWCD_CY` to create a list.")
} else {
  cat("`cons_NonCAWCD_CY` does not have at least two columns as expected.")
}
```

* Total CAWCD System Conservation of `r sprintf("%.1f",round(sum(cons_CAWCD[,column_number])/1000,1))` kaf

```{r generate-bullet-list-CAWCD-Y1, results='asis'}
# Check if cons_CA_CY has rows and at least two columns before proceeding
if (nrow(cons_CAWCD_CY) > 0 && ncol(cons_CAWCD_CY) >= 2) {

  # Use apply to iterate over rows. For each row:
  # row[1] will be the value from the first column of cons_CA_CY
  # row[2] will be the value from the second column of cons_CA_CY
  bullet_points <- apply(cons_CAWCD_CY, 1, function(row) {
    paste0("  * ", row[1], " : ", ifelse(as.numeric(row[2])>100, paste0(round(((as.numeric(row[2])/1000)*100+.1),-1)/100 , " kaf\n"), paste0(round((as.numeric(row[2])*10+.1),-1)/10, " af\n") ))
  })
  
  # Print each bullet point, separated by a newline
  cat(paste0(bullet_points, collapse = ""))
  
} else if (nrow(cons_CAWCD_CY) == 0) {
  cat("No data available in `cons_CAWCD_CY` to create a list.")
} else {
  cat("`cons_CAWCD_CY` does not have at least two columns as expected.")
}


```

```{r conditional-AZ-BHC-Y1, results='asis', echo=FALSE}

if((as.numeric(find_rows_containing_string(DataY1, "BHC")$value)*-1)>0){
  addBHC = TRUE
  volBHC = format(as.numeric(find_rows_containing_string(DataY1, "BHC")$value)*-1, big.mark = ",")
} else{
  addBHC = FALSE
  volBHC=""
}

if (exists("addBHC") && addBHC) {
  
  cat("* Bullhead City PSCP volume of ")
  cat(volBHC)
  cat(" af\n")
}

```

```{r conditional-AZ-242-Y1, results='asis', echo=FALSE}

if(find_rows_containing_string(cons_NonCAWCD,"242 Wellfield")[,column_number]>0){
  add242 = TRUE
  vol242 = sprintf("%.1f",round(find_rows_containing_string(cons_NonCAWCD,"242 Wellfield")[,column_number]/1000,1))
} else{
  add242 = FALSE
  vol242=""
}

if (exists("add242") && add242) {
  
  cat("* System water created by the 242 Well Field Expansion Project of ")
  cat(vol242)
  cat(" kaf\n")
}

```


```{r conditional-AZ-BICS-Y1, results='asis', echo=FALSE}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by CAWCD\n")
}

```

##### Nevada: [`r sprintf("%.3f",round(find_rows_containing_string(sct_data, "NvTotalAnnual")[1,2]/1000000,3))` maf]{style="color: #0070C0;"}
* SNWA annual use of `r round(find_rows_containing_string(sct_data, "SNWPAnnualFC")[1,2]/1000,0)` kaf. Projected diversion includes:

```{r conditional-NV-Shortage, results='asis', echo=FALSE}
# This R code chunk will print Markdown text if the condition is met.
# The 'results="asis"' option is crucial for this to work.

if(find_rows_containing_string(sct_data, "SNWP Annual Shortage Volume")[1,2]>0){
  addNVShortage = TRUE
  nvshort = unlist(unname(round(find_rows_containing_string(sct_data, "SNWP Annual Shortage Volume")[1,2]/1000,0)))
} else{
  addNVShortage = FALSE
  nvshort=""
}

if(find_rows_containing_string(sct_data, "VacatedECICSToBeDelivered_SNWA")[1,2]>0){
  addECICSdel = TRUE
  nvECICSdel = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedECICSToBeDelivered_SNWA")[1,2]/1000,1)))
} else{
  addECICSdel = FALSE
  nvECICSdel=""
}

if (exists("addNVShortage") && addNVShortage) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Shortage volume of ")
  cat(nvshort)
  cat(" kaf\n")
}

if (exists("addECICSdel") && addECICSdel) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Delivery of ")
  cat(nvECICSdel)
  cat(" kaf of EC ICS due to full ICS bank\n")
}

```


```{r conditional-NV-DCP, results='asis', echo=FALSE}
# This R code chunk will print Markdown text if the condition is met.
# The 'results="asis"' option is crucial for this to work.

if(find_rows_containing_string(sct_data, "AnnualCreationDCP_NV")[1,2]>0){
  addNVDCP = TRUE
  nvdcp = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationDCP_NV")[1,2]/1000,0)))
} else{
  addNVDCP = FALSE
  nvdcp=""
}

if (exists("addNVDCP") && addNVDCP) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("* DCP contribution of ")
  cat(nvdcp)
  cat(" kaf through EC ICS conversion\n")
}


```


```{r conditional-NV-SysCons, results='asis', echo=FALSE}
# This R code chunk will print Markdown text if the condition is met.
# The 'results="asis"' option is crucial for this to work.

if(find_rows_containing_string(sct_data, "NV_SystemConservation")[1,2]>0){
  addNVTotSyscon = TRUE
  nvTotSyscon = unlist(unname(round(find_rows_containing_string(sct_data, "NV_SystemConservation")[1,2]/1000,0)))
} else{
  addNVTotSyscon = FALSE
  nvTotSyscon=""
}

if(find_rows_containing_string(sct_data, "NV_LeftinMead")[1,2]>0){
  addTribSyscon = TRUE
  nvTribSyscon = unlist(unname(round(find_rows_containing_string(sct_data, "NV_LeftinMead")[1,2]/1000,0)))
} else{
  addTribSyscon = FALSE
  nvTribSyscon=""
}


if (exists("addNVTotSyscon") && addNVTotSyscon) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("* Total System Conservation of ")
  cat(nvTotSyscon)
  cat(" kaf\n")
}

if (exists("addTribSyscon") && addTribSyscon) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Tributary conservation of ")
  cat(nvTribSyscon)
  cat(" kaf\n")
}

if(addNVTotSyscon && addTribSyscon) {
  if((nvTotSyscon - nvTribSyscon)>0){
    cat("  * Other system conservation of ")
    cat(nvTotSyscon - nvTribSyscon)
    cat(" kaf\n")
  }
}


```

```{r conditional-NV-BICS-Y1, results='asis', echo=FALSE}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by SNWA\n")
}

if(find_rows_containing_string(sct_data, "VacatedDCPICSToSysWater_SNWA")[1,2]>0){
  addVacDCPtoSys = TRUE
  nvVacDCPtoSys = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedDCPICSToSysWater_SNWA")[1,2]/1000,1)))
} else{
  addVacDCPtoSys = FALSE
  nvVacDCPtoSys=""
}

if(find_rows_containing_string(sct_data, "VacatedECICSToSysWater_SNWA")[1,2]>0){
  addVacECICStoSys = TRUE
  nvVacECICStoSys = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedECICSToSysWater_SNWA")[1,2]/1000,1)))
} else{
  addVacECICStoSys = FALSE
  nvVacECICStoSys=""
}

if (addVacDCPtoSys && addVacECICStoSys) {
  # Both are TRUE: assign the sum
  totaltosys <- nvVacDCPtoSys + nvVacECICStoSys
} else if (addVacDCPtoSys) {
  # Only addVacDCPtoSys is TRUE: assign its corresponding numeric value
  totaltosys <- nvVacDCPtoSys
} else if (addVacECICStoSys) {
  # Only addVacECICStoSys is TRUE: assign its corresponding numeric value
  totaltosys <- nvVacECICStoSys # Using the assumed numeric variable name
}

if (addVacDCPtoSys || addVacECICStoSys) {
  cat("* Vacated ICS space (to system water) of ") 
  cat(totaltosys)
  cat(" kaf\n")
}

if (addVacDCPtoSys){
  cat("  * DCP ICS of ")
  cat(nvVacDCPtoSys)
  cat(" kaf converted to system water\n")
}

if (addVacECICStoSys){
  cat("  * EC ICS of ")
  cat(nvVacECICStoSys)
  cat(" kaf converted to system water\n")
}

```

#### Mexico’s Scheduled Water Delivery: `r sprintf("%.3f", round(as.numeric(find_rows_containing_string(DataY1, "Mexico Use")$value)/1000000,3))` maf

```{r conditional-Mx-short-Y1, results='asis', echo=FALSE}
# This R code chunk will print Markdown text if the condition is met.
# The 'results="asis"' option is crucial for this to work.

if(find_rows_containing_string(sct_data, "Mexico Annual Shortage")[1,2]>0){
  addMxShort = TRUE
  mxShort = unlist(unname(round(find_rows_containing_string(sct_data, "Mexico Annual Shortage")[1,2]/1000,0)))
} else{
  addMxShort = FALSE
  mxShort=""
}

if(find_rows_containing_string(sct_data, "AnnualCreationMXRecoverableWaterSavings")[1,2]>0){
  createWaterSav = TRUE
  mxCreateWaterSav = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationMXRecoverableWaterSavings")[1,2]/1000,0)))
} else{
  createWaterSav = FALSE
  mxCreateWaterSav=""
}

if(find_rows_containing_string(sct_data, "AnnualDeliveryMXRecoverableWaterSavings")[1,2]>0){
  delWaterSav = TRUE
  mxdelWaterSav = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDeliveryMXRecoverableWaterSavings")[1,2]/1000,0)))
} else{
  delWaterSav = FALSE
  mxdelWaterSav=""
}

if(find_rows_containing_string(sct_data, "MX_SystemConservation")[1,2]>0){
  createsysCons = TRUE
  mxSysCon = unlist(unname(round(find_rows_containing_string(sct_data, "MX_SystemConservation")[1,2]/1000,1)))
} else{
  createsysCons = FALSE
  mxSysCon=""
}

if(find_rows_containing_string(sct_data, "AnnualDeliveryMexicoWaterReserve")[1,2]>0){
  delMWR = TRUE
  mxdelMWR = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDeliveryMexicoWaterReserve")[1,2]/1000,1)))
} else{
  delMWR = FALSE
  mxdelMWR=""
}

if(find_rows_containing_string(sct_data, "AnnualCreationMexicoWaterReserve")[1,2]>0){
  createMWR = TRUE
  mxcreateMWR = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationMexicoWaterReserve")[1,2]/1000,1)))
} else{
  createMWR = FALSE
  mxcreateMWR=""
}

if (exists("createWaterSav") && createWaterSav) {
  cat("* Projected delivery includes:\n")
  
  if (exists("addMxShort") && addMxShort) {
    cat("  * Shortage volume of ")
    cat(mxShort)
    cat(" kaf\n")
  }
  
  cat("  * Recoverable Water Savings Contribution of ")
  cat(mxCreateWaterSav)
  cat(" kaf\n")
  
  if (exists("delWaterSav") && delWaterSav) {
    cat("  * Recoverable Water Savings Delivery of ")
    cat(mxdelWaterSav)
    cat(" kaf\n")
  }
  
  if (exists("createsysCons") && createsysCons) {
    cat("  * Minute 330 System Conservation of ")
    cat(mxSysCon)
    cat(" kaf\n")
  }
  
  if (exists("delMWR") && delMWR) {
    cat("* Water Reserve delivery of ")
    cat(mxdelMWR)
    cat(" kaf\n")
  }
  
  if (exists("createMWR") && createMWR) {
    cat("* Water Reserve creation of ")
    cat(mxcreateMWR)
    cat(" kaf\n")
  }
}

```


### `r as.numeric(current_year) + 1`

Total projected water use **(`r sprintf("%.3f",round((find_rows_containing_string(sct_data,"MexicoAdjustedSched")[3] + sum(apportionment[,3]))/1000000,3))` maf)** – `<w:r><w:rPr><w:highlight w:val="yellow"/><w:i/></w:rPr><w:t>Based on Lake Mead Operating Condition of Level 1 Shortage and water savings contributions under the LB DCP Agreement and IBWC Minute 323.</w:t></w:r>`{=openxml}

#### U.S. Contractors: `r sprintf("%.3f",round(sum(apportionment[,3])/1000000,3))` maf

##### California: [`r sprintf("%.3f",round(sum(apportionment[1,3])/1000000,3))` maf]{style="color: #0070C0;"}
* MWD annual diversion of `r prettyNum(round(as.numeric(find_rows_containing_string(sct_data,"MWDDiversionAnnualFC")[3])/1000,0),  big.mark = ",", scientific = FALSE)` kaf
    
```{r nwd-diversion-y2, results='asis'}
cat("  * Projected diversion includes the ")
cat(mwdICSY2)
cat(" kaf of ICS")

if(find_rows_containing_string(sct_data, "AnnualDCPContribution_Coachella")[1,3]>0){
  addCADCP = TRUE
  coachDCP = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDCPContribution_Coachella")[1,3]/1000,1)))
  mwdDCP = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDCPContribution_MWD")[1,3]/1000,1)))
} else{
  addCADCP = FALSE
  coachDCP=""
  mwdDCP=""
}

if (exists("addCADCP") && addCADCP) {
  
  cat(" and ")
  cat(coachDCP)
  cat(" kaf of water stored for CVWD's share of California's DCP contribution\n")
  cat("* DCP contribution of ")
  cat(mwdDCP + coachDCP)
  cat(" kaf through EC ICS conversion")
}

cat("\n")

```


```{r generate-bullet-list-apply-2, results='asis'}
# Check if cons_CA_CY2 has rows and at least two columns before proceeding

if (nrow(cons_CA_CY2) > 0 && ncol(cons_CA_CY2) >= 2) {
  
  cat("* Total California System Conservation of ")
  cat(sprintf("%.1f",round(sum(cons_CA[,column_number+1])/1000,1)))
  cat(" kaf\n")

  # Use apply to iterate over rows. For each row:
  # row[1] will be the value from the first column of cons_CA_CY
  # row[2] will be the value from the second column of cons_CA_CY
  bullet_points <- apply(cons_CA_CY2, 1, function(row) {
    paste0("  * ", row[1], " system conservation of ", round(as.numeric(row[2])/1000,1) , " kaf\n")
  })
  
  # Print each bullet point, separated by a newline
  cat(paste0(bullet_points, collapse = ""))
  
} 

```

* Needles PSCP volume of `r as.numeric(find_rows_containing_string(DataY2, "Needles")$value)*-1` af
```{r conditional-CA-BICS-Y2, results='asis', echo=FALSE}

if(find_rows_containing_string(sct_data, "AnnualCreationBiNat_MWD")[1,3]>0){
  addBICS = TRUE
  bics = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationBiNat_MWD")[1,3]/1000,1)))
} else{
  addBICS = FALSE
  bics=""
}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by MWD and IID\n")
}


```

##### Arizona: [`r sprintf("%.3f",round(sum(apportionment[2,3])/1000000,3))` maf]{style="color: #0070C0;"}
* CAP annual diversion of `r prettyNum(round(find_rows_containing_string(sct_data, "CAPAnnualFC")[1,3]/1000,0),  big.mark = ",", scientific = FALSE)` kaf

```{r conditional-AZ-reductions-Y2, results='asis', echo=FALSE}

addreductions = calculate_AZ_reductions(year=2, sct_data = sct_data)$addreductions
reductions_scen_all = calculate_AZ_reductions(year=2, sct_data = sct_data)$reductions_scen_all

if (exists("addreductions") && addreductions) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("    * Projected diversion includes a ")
  cat(reductions_scen_all)
  cat("\n")
}

az_dcpICS <- find_rows_containing_string(sct_data, "AnnualCreationDCP_CAWCD")[3]
az_dcpICS <- ifelse((az_dcpICS/1000) %% 1 != 0, round(az_dcpICS/1000, 1), round(az_dcpICS/1000, 0))
az_dcpSys <- find_rows_containing_string(sct_data, "AnnualSysWaterforDCP_CAWCD")[3]
az_dcpSys <- ifelse((az_dcpSys/1000) %% 1 != 0, round(az_dcpSys/1000, 1), round(az_dcpSys/1000, 0))

```

* DCP contribution will be made by creating `r ifelse(az_dcpICS > 0, paste0(az_dcpICS, " kaf of ICS and ", az_dcpSys, " kaf of non-ICS water"), paste0(az_dcpSys, " kaf of non-ICS water"))` 
* Total non-CAWCD System Conservation of `r sprintf("%.1f",round(sum(cons_NonCAWCD_filt[,column_number +1])/1000,1))` kaf
```{r generate-bullet-list-nonCAWCD-Y2, results='asis'}
# Check if cons_CA_CY has rows and at least two columns before proceeding
if (nrow(cons_NonCAWCD_CY2) > 0 && ncol(cons_NonCAWCD_CY2) >= 2) {

  # Use apply to iterate over rows. For each row:
  # row[1] will be the value from the first column of cons_CA_CY
  # row[2] will be the value from the second column of cons_CA_CY
  bullet_points <- apply(cons_NonCAWCD_CY2, 1, function(row) {
    paste0("  * ", row[1], " : ", ifelse(as.numeric(row[2])>100, paste0(round(((as.numeric(row[2])/1000)*100+.1),-1)/100 , " kaf\n"), paste0(round((as.numeric(row[2])*10+.1),-1)/10, " af\n") ))
  })
  
  # Print each bullet point, separated by a newline
  cat(paste0(bullet_points, collapse = ""))
  
} 

```

* Total CAWCD System Conservation of `r sprintf("%.1f",round(sum(cons_CAWCD[,column_number +1])/1000,1))` kaf

```{r generate-bullet-list-CAWCD-Y2, results='asis'}
# Check if cons_CA_CY has rows and at least two columns before proceeding
if (nrow(cons_CAWCD_CY2) > 0 && ncol(cons_CAWCD_CY2) >= 2) {

  # Use apply to iterate over rows. For each row:
  # row[1] will be the value from the first column of cons_CA_CY
  # row[2] will be the value from the second column of cons_CA_CY
  bullet_points <- apply(cons_CAWCD_CY2, 1, function(row) {
    paste0("  * ", row[1], " : ", ifelse(as.numeric(row[2])>100, paste0(round(((as.numeric(row[2])/1000)*100+.1),-1)/100 , " kaf\n"), paste0(round((as.numeric(row[2])*10+.1),-1)/10, " af\n") ))
  })
  
  # Print each bullet point, separated by a newline
  cat(paste0(bullet_points, collapse = ""))
  
} 

```

```{r conditional-AZ-BHC-Y2, results='asis', echo=FALSE}

if((as.numeric(find_rows_containing_string(DataY2, "BHC")$value)*-1)>0){
  addBHC = TRUE
  volBHC = format(as.numeric(find_rows_containing_string(DataY2, "BHC")$value)*-1, big.mark = ",")
} else{
  addBHC = FALSE
  volBHC=""
}

if (exists("addBHC") && addBHC) {
  
  cat("* Bullhead City PSCP volume of ")
  cat(volBHC)
  cat(" af\n")
}

```

```{r conditional-AZ-242-Y2, results='asis', echo=FALSE}

if(find_rows_containing_string(cons_NonCAWCD,"242 Wellfield")[,column_number+1]>0){
  add242 = TRUE
  vol242 = sprintf("%.1f",round(find_rows_containing_string(cons_NonCAWCD,"242 Wellfield")[,column_number+1]/1000,1))
} else{
  add242 = FALSE
  vol242=""
}

if (exists("add242") && add242) {
  
  cat("* System water created by the 242 Well Field Expansion Project of ")
  cat(vol242)
  cat(" kaf\n")
}

```


```{r conditional-AZ-BICS-Y2, results='asis', echo=FALSE}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by CAWCD\n")
}

```

##### Nevada: [`r sprintf("%.3f",round(find_rows_containing_string(sct_data, "NvTotalAnnual")[1,3]/1000000,3))` maf]{style="color: #0070C0;"}

* SNWA annual use of `r round(find_rows_containing_string(sct_data, "SNWPAnnualFC")[1,3]/1000,0)` kaf. Projected diversion includes:

```{r conditional-NV-Shortage-Y2, results='asis', echo=FALSE}
# This R code chunk will print Markdown text if the condition is met.
# The 'results="asis"' option is crucial for this to work.

if(find_rows_containing_string(sct_data, "SNWP Annual Shortage Volume")[1,3]>0){
  addNVShortage = TRUE
  nvshort = unlist(unname(round(find_rows_containing_string(sct_data, "SNWP Annual Shortage Volume")[1,3]/1000,0)))
} else{
  addNVShortage = FALSE
  nvshort=""
}

if(find_rows_containing_string(sct_data, "VacatedECICSToBeDelivered_SNWA")[1,3]>0){
  addECICSdel = TRUE
  nvECICSdel = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedECICSToBeDelivered_SNWA")[1,3]/1000,1)))
} else{
  addECICSdel = FALSE
  nvECICSdel=""
}

if (exists("addNVShortage") && addNVShortage) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Shortage volume of ")
  cat(nvshort)
  cat(" kaf\n")
}

if (exists("addECICSdel") && addECICSdel) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Delivery of ")
  cat(nvECICSdel)
  cat(" kaf of EC ICS due to full ICS bank\n")
}

```


```{r conditional-NV-DCP-Y2, results='asis', echo=FALSE}
# This R code chunk will print Markdown text if the condition is met.
# The 'results="asis"' option is crucial for this to work.

if(find_rows_containing_string(sct_data, "AnnualCreationDCP_NV")[1,3]>0){
  addNVDCP = TRUE
  nvdcp = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationDCP_NV")[1,3]/1000,0)))
} else{
  addNVDCP = FALSE
  nvdcp=""
}

if (exists("addNVDCP") && addNVDCP) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("* DCP contribution of ")
  cat(nvdcp)
  cat(" kaf through EC ICS conversion\n")
}


```


```{r conditional-NV-SysCons-Y2, results='asis', echo=FALSE}
# This R code chunk will print Markdown text if the condition is met.
# The 'results="asis"' option is crucial for this to work.

if(find_rows_containing_string(sct_data, "NV_SystemConservation")[1,3]>0){
  addNVTotSyscon = TRUE
  nvTotSyscon = unlist(unname(round(find_rows_containing_string(sct_data, "NV_SystemConservation")[1,3]/1000,0)))
} else{
  addNVTotSyscon = FALSE
  nvTotSyscon=""
}

if(find_rows_containing_string(sct_data, "NV_LeftinMead")[1,3]>0){
  addTribSyscon = TRUE
  nvTribSyscon = unlist(unname(round(find_rows_containing_string(sct_data, "NV_LeftinMead")[1,3]/1000,0)))
} else{
  addTribSyscon = FALSE
  nvTribSyscon=""
}


if (exists("addNVTotSyscon") && addNVTotSyscon) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("* Total System Conservation of ")
  cat(nvTotSyscon)
  cat(" kaf\n")
}

if (exists("addTribSyscon") && addTribSyscon) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Tributary conservation of ")
  cat(nvTribSyscon)
  cat(" kaf\n")
}

if(addNVTotSyscon && addTribSyscon) {
  if((nvTotSyscon - nvTribSyscon)>0){
    cat("  * Other system conservation of ")
    cat(nvTotSyscon - nvTribSyscon)
    cat(" kaf\n")
  }
}


```

```{r conditional-NV-BICS-Y2, results='asis', echo=FALSE}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by SNWA\n")
}

if(find_rows_containing_string(sct_data, "VacatedDCPICSToSysWater_SNWA")[1,3]>0){
  addVacDCPtoSys = TRUE
  nvVacDCPtoSys = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedDCPICSToSysWater_SNWA")[1,3]/1000,1)))
} else{
  addVacDCPtoSys = FALSE
  nvVacDCPtoSys=""
}

if(find_rows_containing_string(sct_data, "VacatedECICSToSysWater_SNWA")[1,3]>0){
  addVacECICStoSys = TRUE
  nvVacECICStoSys = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedECICSToSysWater_SNWA")[1,3]/1000,1)))
} else{
  addVacECICStoSys = FALSE
  nvVacECICStoSys=""
}

if (addVacDCPtoSys && addVacECICStoSys) {
  # Both are TRUE: assign the sum
  totaltosys <- nvVacDCPtoSys + nvVacECICStoSys
} else if (addVacDCPtoSys) {
  # Only addVacDCPtoSys is TRUE: assign its corresponding numeric value
  totaltosys <- nvVacDCPtoSys
} else if (addVacECICStoSys) {
  # Only addVacECICStoSys is TRUE: assign its corresponding numeric value
  totaltosys <- nvVacECICStoSys # Using the assumed numeric variable name
}

if (addVacDCPtoSys || addVacECICStoSys) {
  cat("* Vacated ICS space (to system water) of ") 
  cat(totaltosys)
  cat(" kaf\n")
}

if (addVacDCPtoSys){
  cat("  * DCP ICS of ")
  cat(nvVacDCPtoSys)
  cat(" kaf converted to system water\n")
}

if (addVacECICStoSys){
  cat("  * EC ICS of ")
  cat(nvVacECICStoSys)
  cat(" kaf converted to system water\n")
}

```


#### Mexico’s Scheduled Water Delivery: `r sprintf("%.3f", round(find_rows_containing_string(sct_data,"MexicoAdjustedSched")[3]/1000000,3))` maf

```{r conditional-Mx-short-Y2, results='asis', echo=FALSE}
# This R code chunk will print Markdown text if the condition is met.
# The 'results="asis"' option is crucial for this to work.

if(find_rows_containing_string(sct_data, "Mexico Annual Shortage")[1,3]>0){
  addMxShort = TRUE
  mxShort = unlist(unname(round(find_rows_containing_string(sct_data, "Mexico Annual Shortage")[1,3]/1000,0)))
} else{
  addMxShort = FALSE
  mxShort=""
}

if(find_rows_containing_string(sct_data, "AnnualCreationMXRecoverableWaterSavings")[1,3]>0){
  createWaterSav = TRUE
  mxCreateWaterSav = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationMXRecoverableWaterSavings")[1,3]/1000,0)))
} else{
  createWaterSav = FALSE
  mxCreateWaterSav=""
}

if(find_rows_containing_string(sct_data, "AnnualDeliveryMXRecoverableWaterSavings")[1,3]>0){
  delWaterSav = TRUE
  mxdelWaterSav = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDeliveryMXRecoverableWaterSavings")[1,3]/1000,0)))
} else{
  delWaterSav = FALSE
  mxdelWaterSav=""
}

if(find_rows_containing_string(sct_data, "MX_SystemConservation")[1,3]>0){
  createsysCons = TRUE
  mxSysCon = unlist(unname(round(find_rows_containing_string(sct_data, "MX_SystemConservation")[1,3]/1000,1)))
} else{
  createsysCons = FALSE
  mxSysCon=""
}

if(find_rows_containing_string(sct_data, "AnnualDeliveryMexicoWaterReserve")[1,3]>0){
  delMWR = TRUE
  mxdelMWR = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDeliveryMexicoWaterReserve")[1,3]/1000,1)))
} else{
  delMWR = FALSE
  mxdelMWR=""
}

if(find_rows_containing_string(sct_data, "AnnualCreationMexicoWaterReserve")[1,3]>0){
  createMWR = TRUE
  mxcreateMWR = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationMexicoWaterReserve")[1,3]/1000,1)))
} else{
  createMWR = FALSE
  mxcreateMWR=""
}

if (exists("createWaterSav") && createWaterSav) {
  cat("* Projected delivery includes:\n")
  
  if (exists("addMxShort") && addMxShort) {
    cat("  * Shortage volume of ")
    cat(mxShort)
    cat(" kaf\n")
  }
  
  cat("  * Recoverable Water Savings Contribution of ")
  cat(mxCreateWaterSav)
  cat(" kaf\n")
  
  if (exists("delWaterSav") && delWaterSav) {
    cat("  * Recoverable Water Savings Delivery of ")
    cat(mxdelWaterSav)
    cat(" kaf\n")
  }
  
  if (exists("createsysCons") && createsysCons) {
    cat("  * Minute 330 System Conservation of ")
    cat(mxSysCon)
    cat(" kaf\n")
  }
  
  if (exists("delMWR") && delMWR) {
    cat("* Water Reserve delivery of ")
    cat(mxdelMWR)
    cat(" kaf\n")
  }
  
  if (exists("createMWR") && createMWR) {
    cat("* Water Reserve creation of ")
    cat(mxcreateMWR)
    cat(" kaf\n")
  }
}

```


### `r as.numeric(current_year) + 2`

Total projected water use **(`r sprintf("%.3f",round((find_rows_containing_string(sct_data,"MexicoAdjustedSched")[4] + sum(apportionment[,4]))/1000000,3))` maf)** – `<w:r><w:rPr><w:highlight w:val="yellow"/><w:i/></w:rPr><w:t>Based on Lake Mead Operating Condition of Level 2 Shortage and water savings contributions under the LB DCP Agreement and IBWC Minute 323. For modeling purposes, simulated years beyond 2026 assume a continuation of the 2007 Interim Guidelines, the 2019 Colorado River Basin Drought Contingency Plans, and Minute 323, including the Binational Water Scarcity Contingency Plan. Except for certain provisions related to ICS recovery and Upper Basin demand management, operations under these agreements are in effect through 2026. Reclamation initiated the process to develop operations for post-2026 in June 2023, and the modeling assumptions described here are subject to change.</w:t></w:r>`{=openxml}

#### U.S. Contractors: `r sprintf("%.3f",round(sum(apportionment[,4])/1000000,3))` maf

##### California: [`r sprintf("%.3f",round(sum(apportionment[1,4])/1000000,3))` maf]{style="color: #0070C0;"}
* MWD annual diversion of `r prettyNum(round(as.numeric(find_rows_containing_string(sct_data,"MWDDiversionAnnualFC")[4])/1000,0),  big.mark = ",", scientific = FALSE)` kaf
    
```{r nwd-diversion-y3, results='asis'}
cat("  * Projected diversion includes the ")
cat(mwdICSY3)
cat(" kaf of ICS")

if(find_rows_containing_string(sct_data, "AnnualDCPContribution_Coachella")[1,4]>0){
  addCADCP = TRUE
  coachDCP = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDCPContribution_Coachella")[1,4]/1000,1)))
  mwdDCP = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDCPContribution_MWD")[1,4]/1000,1)))
} else{
  addCADCP = FALSE
  coachDCP=""
  mwdDCP=""
}

if (exists("addCADCP") && addCADCP) {
  
  cat(" and ")
  cat(coachDCP)
  cat(" kaf of water stored for CVWD's share of California's DCP contribution\n")
  cat("* DCP contribution of ")
  cat(mwdDCP + coachDCP)
  cat(" kaf through EC ICS conversion")
}

cat("\n")

```

```{r generate-bullet-list-apply-3, results='asis'}

if (nrow(cons_CA_CY3) > 0 && ncol(cons_CA_CY3) >= 2) {
  
  cat("* Total California System Conservation of ")
  cat(sprintf("%.1f",round(sum(cons_CA[,column_number+2])/1000,1)))
  cat(" kaf\n")

  bullet_points <- apply(cons_CA_CY3, 1, function(row) {
    paste0("  * ", row[1], " system conservation of ", round(as.numeric(row[2])/1000,1) , " kaf\n")
  })
  
  cat(paste0(bullet_points, collapse = ""))
} 

```

* Needles PSCP volume of `r as.numeric(find_rows_containing_string(DataY2, "Needles")$value)*-1` af

```{r conditional-CA-BICS-Y3, results='asis', echo=FALSE}

if(find_rows_containing_string(sct_data, "AnnualCreationBiNat_MWD")[1,4]>0){
  addBICS = TRUE
  bics = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationBiNat_MWD")[1,4]/1000,1)))
} else{
  addBICS = FALSE
  bics=""
}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by MWD and IID\n")
}

```

##### Arizona: [`r sprintf("%.3f",round(sum(apportionment[2,4])/1000000,3))` maf]{style="color: #0070C0;"}
* CAP annual diversion of `r prettyNum(round(find_rows_containing_string(sct_data, "CAPAnnualFC")[1,4]/1000,0),  big.mark = ",", scientific = FALSE)` kaf

```{r conditional-AZ-reductions-Y3, results='asis', echo=FALSE}

addreductions = calculate_AZ_reductions(year=3, sct_data = sct_data)$addreductions
reductions_scen_all = calculate_AZ_reductions(year=3, sct_data = sct_data)$reductions_scen_all

if (exists("addreductions") && addreductions) {
  cat("    * Projected diversion includes a ")
  cat(reductions_scen_all)
  cat("\n")
}

az_dcpICS <- find_rows_containing_string(sct_data, "AnnualCreationDCP_CAWCD")[4]
az_dcpICS <- ifelse((az_dcpICS/1000) %% 1 != 0, round(az_dcpICS/1000, 1), round(az_dcpICS/1000, 0))
az_dcpSys <- find_rows_containing_string(sct_data, "AnnualSysWaterforDCP_CAWCD")[4]
az_dcpSys <- ifelse((az_dcpSys/1000) %% 1 != 0, round(az_dcpSys/1000, 1), round(az_dcpSys/1000, 0))

```

* DCP contribution will be made by creating `r ifelse(az_dcpICS > 0, paste0(az_dcpICS, " kaf of ICS and ", az_dcpSys, " kaf of non-ICS water"), paste0(az_dcpSys, " kaf of non-ICS water"))` 
* Total non-CAWCD System Conservation of `r sprintf("%.1f",round(sum(cons_NonCAWCD_filt[,column_number +2])/1000,1))` kaf

```{r generate-bullet-list-nonCAWCD-Y3, results='asis'}

if (nrow(cons_NonCAWCD_CY3) > 0 && ncol(cons_NonCAWCD_CY3) >= 2) {

  bullet_points <- apply(cons_NonCAWCD_CY3, 1, function(row) {
    paste0("  * ", row[1], " : ", ifelse(as.numeric(row[2])>100, paste0(round(((as.numeric(row[2])/1000)*100+.1),-1)/100 , " kaf\n"), paste0(round((as.numeric(row[2])*10+.1),-1)/10, " af\n") ))
  })
  
  cat(paste0(bullet_points, collapse = ""))
} 

```

* Total CAWCD System Conservation of `r sprintf("%.1f",round(sum(cons_CAWCD[,column_number +2])/1000,1))` kaf

```{r generate-bullet-list-CAWCD-Y3, results='asis'}

if (nrow(cons_CAWCD_CY3) > 0 && ncol(cons_CAWCD_CY3) >= 2) {

  bullet_points <- apply(cons_CAWCD_CY3, 1, function(row) {
    paste0("  * ", row[1], " : ", ifelse(as.numeric(row[2])>100, paste0(round(((as.numeric(row[2])/1000)*100+.1),-1)/100 , " kaf\n"), paste0(round((as.numeric(row[2])*10+.1),-1)/10, " af\n") ))
  })
  
  cat(paste0(bullet_points, collapse = ""))
} 

```


```{r conditional-AZ-BHC-Y3, results='asis', echo=FALSE}

if((as.numeric(find_rows_containing_string(DataY3, "BHC")$value)*-1)>0){
  addBHC = TRUE
  volBHC = format(as.numeric(find_rows_containing_string(DataY3, "BHC")$value)*-1, big.mark = ",")
} else{
  addBHC = FALSE
  volBHC=""
}

if (exists("addBHC") && addBHC) {
  
  cat("* Bullhead City PSCP volume of ")
  cat(volBHC)
  cat(" af\n")
}

```


```{r conditional-AZ-242-Y3, results='asis', echo=FALSE}

if(find_rows_containing_string(cons_NonCAWCD,"242 Wellfield")[,column_number+2]>0){
  add242 = TRUE
  vol242 = sprintf("%.1f",round(find_rows_containing_string(cons_NonCAWCD,"242 Wellfield")[,column_number+2]/1000,1))
} else{
  add242 = FALSE
  vol242=""
}

if (exists("add242") && add242) {
  
  cat("* System water created by the 242 Well Field Expansion Project of ")
  cat(vol242)
  cat(" kaf\n")
}

```


```{r conditional-AZ-BICS-Y3, results='asis', echo=FALSE}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by CAWCD\n")
}

```


##### Nevada: [`r sprintf("%.3f",round(find_rows_containing_string(sct_data, "NvTotalAnnual")[1,4]/1000000,3))` maf]{style="color: #0070C0;"}

* SNWA annual use of `r round(find_rows_containing_string(sct_data, "SNWPAnnualFC")[1,4]/1000,0)` kaf. Projected diversion includes:

```{r conditional-NV-Shortage-Y3, results='asis', echo=FALSE}

if(find_rows_containing_string(sct_data, "SNWP Annual Shortage Volume")[1,4]>0){
  addNVShortage = TRUE
  nvshort = unlist(unname(round(find_rows_containing_string(sct_data, "SNWP Annual Shortage Volume")[1,4]/1000,0)))
} else{
  addNVShortage = FALSE
  nvshort=""
}

if(find_rows_containing_string(sct_data, "AnnualCreationEC_NV")[1,4]>0){
  addECICScreate = TRUE
  nvECICSCreate = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationEC_NV")[1,4]/1000,1)))
} else{
  addECICScreate = FALSE
  nvECICSCreate=""
}

if(find_rows_containing_string(sct_data, "VacatedECICSToBeDelivered_SNWA")[1,4]>0){
  addECICSdel = TRUE
  nvECICSdel = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedECICSToBeDelivered_SNWA")[1,4]/1000,1)))
} else{
  addECICSdel = FALSE
  nvECICSdel=""
}

if (exists("addNVShortage") && addNVShortage) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Shortage volume of ")
  cat(nvshort)
  cat(" kaf\n")
}

if (exists("addECICScreate") && addECICScreate) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * EC ICS creation of ")
  cat(nvECICSCreate)
  cat(" kaf\n")
}

if (exists("addECICSdel") && addECICSdel) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Delivery of ")
  cat(nvECICSdel)
  cat(" kaf of EC ICS due to full ICS bank\n")
}


```


```{r conditional-NV-DCP-Y3, results='asis', echo=FALSE}

if(find_rows_containing_string(sct_data, "AnnualCreationDCP_NV")[1,4]>0){
  addNVDCP = TRUE
  nvdcp = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationDCP_NV")[1,4]/1000,0)))
} else{
  addNVDCP = FALSE
  nvdcp=""
}

if (exists("addNVDCP") && addNVDCP) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("* DCP contribution of ")
  cat(nvdcp)
  cat(" kaf through EC ICS conversion\n")
}


```


```{r conditional-NV-SysCons-Y3, results='asis', echo=FALSE}

if(find_rows_containing_string(sct_data, "NV_SystemConservation")[1,4]>0){
  addNVTotSyscon = TRUE
  nvTotSyscon = unlist(unname(round(find_rows_containing_string(sct_data, "NV_SystemConservation")[1,4]/1000,0)))
} else{
  addNVTotSyscon = FALSE
  nvTotSyscon=""
}

if(find_rows_containing_string(sct_data, "NV_LeftinMead")[1,4]>0){
  addTribSyscon = TRUE
  nvTribSyscon = unlist(unname(round(find_rows_containing_string(sct_data, "NV_LeftinMead")[1,4]/1000,0)))
} else{
  addTribSyscon = FALSE
  nvTribSyscon=""
}


if (exists("addNVTotSyscon") && addNVTotSyscon) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("* Total System Conservation of ")
  cat(nvTotSyscon)
  cat(" kaf\n")
}

if (exists("addTribSyscon") && addTribSyscon) {
  #cat("\n") # Ensures a new line before the conditional text
  cat("  * Tributary conservation of ")
  cat(nvTribSyscon)
  cat(" kaf\n")
}

if(addNVTotSyscon && addTribSyscon) {
  if((nvTotSyscon - nvTribSyscon)>0){
    cat("  * Other system conservation of ")
    cat(nvTotSyscon - nvTribSyscon)
    cat(" kaf\n")
  }
}

```

```{r conditional-NV-BICS-Y3, results='asis', echo=FALSE}

if (exists("addBICS") && addBICS) {
  
  cat("* Binational ICS creation of ")
  cat(bics)
  cat(" kaf by SNWA\n")
}

if(find_rows_containing_string(sct_data, "VacatedDCPICSToSysWater_SNWA")[1,4]>0){
  addVacDCPtoSys = TRUE
  nvVacDCPtoSys = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedDCPICSToSysWater_SNWA")[1,4]/1000,1)))
} else{
  addVacDCPtoSys = FALSE
  nvVacDCPtoSys=""
}

if(find_rows_containing_string(sct_data, "VacatedECICSToSysWater_SNWA")[1,4]>0){
  addVacECICStoSys = TRUE
  nvVacECICStoSys = unlist(unname(round(find_rows_containing_string(sct_data, "VacatedECICSToSysWater_SNWA")[1,4]/1000,1)))
} else{
  addVacECICStoSys = FALSE
  nvVacECICStoSys=""
}

if (addVacDCPtoSys && addVacECICStoSys) {
  # Both are TRUE: assign the sum
  totaltosys <- nvVacDCPtoSys + nvVacECICStoSys
} else if (addVacDCPtoSys) {
  # Only addVacDCPtoSys is TRUE: assign its corresponding numeric value
  totaltosys <- nvVacDCPtoSys
} else if (addVacECICStoSys) {
  # Only addVacECICStoSys is TRUE: assign its corresponding numeric value
  totaltosys <- nvVacECICStoSys # Using the assumed numeric variable name
}

if (addVacDCPtoSys || addVacECICStoSys) {
  cat("* Vacated ICS space (to system water) of ") 
  cat(totaltosys)
  cat(" kaf\n")
}

if (addVacDCPtoSys){
  cat("  * DCP ICS of ")
  cat(nvVacDCPtoSys)
  cat(" kaf converted to system water\n")
}

if (addVacECICStoSys){
  cat("  * EC ICS of ")
  cat(nvVacECICStoSys)
  cat(" kaf converted to system water\n")
}

```



#### Mexico’s Scheduled Water Delivery: `r sprintf("%.3f", round(find_rows_containing_string(sct_data,"MexicoAdjustedSched")[4]/1000000,3))` maf

```{r conditional-Mx-short-Y3, results='asis', echo=FALSE}

if(find_rows_containing_string(sct_data, "Mexico Annual Shortage")[1,4]>0){
  addMxShort = TRUE
  mxShort = unlist(unname(round(find_rows_containing_string(sct_data, "Mexico Annual Shortage")[1,4]/1000,0)))
} else{
  addMxShort = FALSE
  mxShort=""
}

if(find_rows_containing_string(sct_data, "AnnualCreationMXRecoverableWaterSavings")[1,4]>0){
  createWaterSav = TRUE
  mxCreateWaterSav = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationMXRecoverableWaterSavings")[1,4]/1000,0)))
} else{
  createWaterSav = FALSE
  mxCreateWaterSav=""
}

if(find_rows_containing_string(sct_data, "AnnualDeliveryMXRecoverableWaterSavings")[1,4]>0){
  delWaterSav = TRUE
  mxdelWaterSav = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDeliveryMXRecoverableWaterSavings")[1,4]/1000,0)))
} else{
  delWaterSav = FALSE
  mxdelWaterSav=""
}

if(find_rows_containing_string(sct_data, "MX_SystemConservation")[1,4]>0){
  createsysCons = TRUE
  mxSysCon = unlist(unname(round(find_rows_containing_string(sct_data, "MX_SystemConservation")[1,4]/1000,1)))
} else{
  createsysCons = FALSE
  mxSysCon=""
}

if(find_rows_containing_string(sct_data, "AnnualDeliveryMexicoWaterReserve")[1,4]>0){
  delMWR = TRUE
  mxdelMWR = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualDeliveryMexicoWaterReserve")[1,4]/1000,1)))
} else{
  delMWR = FALSE
  mxdelMWR=""
}

if(find_rows_containing_string(sct_data, "AnnualCreationMexicoWaterReserve")[1,4]>0){
  createMWR = TRUE
  mxcreateMWR = unlist(unname(round(find_rows_containing_string(sct_data, "AnnualCreationMexicoWaterReserve")[1,4]/1000,1)))
} else{
  createMWR = FALSE
  mxcreateMWR=""
}

if (exists("createWaterSav") && createWaterSav) {
  cat("* Projected delivery includes:\n")
  
  if (exists("addMxShort") && addMxShort) {
    cat("  * Shortage volume of ")
    cat(mxShort)
    cat(" kaf\n")
  }
  
  cat("  * Recoverable Water Savings Contribution of ")
  cat(mxCreateWaterSav)
  cat(" kaf\n")
  
  if (exists("delWaterSav") && delWaterSav) {
    cat("  * Recoverable Water Savings Delivery of ")
    cat(mxdelWaterSav)
    cat(" kaf\n")
  }
  
  if (exists("createsysCons") && createsysCons) {
    cat("  * Minute 330 System Conservation of ")
    cat(mxSysCon)
    cat(" kaf\n")
  }
  
  if (exists("delMWR") && delMWR) {
    cat("* Water Reserve delivery of ")
    cat(mxdelMWR)
    cat(" kaf\n")
  }
  
  if (exists("createMWR") && createMWR) {
    cat("* Water Reserve creation of ")
    cat(mxcreateMWR)
    cat(" kaf\n")
  }
}

```


### ICS Totals

* Projected ICS Total Storage at the end of CY `r colnames(transformed_data)[ncol(transformed_data)]`: **`r sprintf("%.3f",transformed_data[nrow(transformed_data),ncol(transformed_data)]/1000000)` maf**
* Projected ICS Storage Balances at the end of each calendar year in the study are as follows:

```{=openxml}
<w:p><w:r><w:br/></w:r></w:p>
```

```{r ics-table}

# --- Display the flextable in Quarto ---
ftics

```

```{=openxml}
<w:p>
  <w:r>
    <w:br w:type="page"/>
  </w:r>
</w:p>
```

### Modeled Lower Basin Conservation Actions
```{r}

# Conservation Table
ft

```

```{=openxml}
<w:p><w:r><w:br/></w:r></w:p>
```

```{r}

# Conservation Summary Table
ft_summary

```
## Notes and Disclaimers

* Modeled Conservation volumes reflect executed agreements and/or current operational projections/assumptions and are subject to change. Additional conservation activities are being considered. After new agreements are finalized and executed, these additional activities will be included in Reclamation’s operational modeling.
* Projected SEIS ROD Reservoir Protection Volume is 3.631 from 2023 through 2026.

